---
title: "Barcode Association Report"
author: "Julia Rühle/Lera Dragan"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    code_folding: hide
    toc_depth: 4
    toc_float:
      collapsed: no
    highlight: pygments
  pdf_document:
    toc: yes
    toc_depth: '4'

  editor_options:
  chunk_output_type: console
params:
  data_file: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_knit$set(root.dir = "..")

require(ggplot2)
require(LSD)
require(plyr)
require(gridExtra)
require(reshape2)
library(patchwork)
library(dplyr)
library(tidyverse)
library(knitr)
library(kableExtra)

# Read the csv file and add 2 columns separating 2 gRNA from the GFP-barcode
all <- read.csv(gzfile(params$data_file), sep = "\t")
all$exp <- "gib2"

all <- all %>%
  separate(BARCODE, into = c("BARCODE", "GUIDE"), sep = "\\|")

# Filter out lines that were not mapped with bwa to GRE or gRNA
mapped <- all %>% 
  filter(REF != "None", GUIDE != "") %>%
  mutate(construct = paste0(REF, GUIDE))

```


# Mapping Stats

```{r mapping stats, include=FALSE}

tot_reads <- sum(all$READS)

perc_mapped <- sum(mapped$READS)/sum(all$READS)

library(knitr)
data <- data.frame(
  Library = "Library",
  "Total No of Reads "= tot_reads, 
  "Mapping Rate total" = perc_mapped
)

kable(data, format = "markdown", align = 'c')

```

# Barcode representation
Most importantly, we want to make sure that unique constructs are represented evenly in the library. We would expect equal number of reads for each construct after sequencing. To look into this, we first compute the number of reads/barcodes per construct. Then we rank each construct, according to its number of reads (Rank 1: The construct with the lowest number of reads).

## How many reads represent one gRNA-barcode combination? 

Each dot represents a unique construct ranked according to the number of reads/barcodes by which it is represented. 

```{r number of reads representing constructs, include = TRUE}
counts_per_construct <-filtered %>%
  group_by(construct) %>%
  summarise(reads_count = sum(READS)) %>%
  mutate(rank = rank(reads_count)) %>%
  arrange(rank)

# Calculate quantiles for READS
quantiles <- quantile(counts_per_construct$reads_count, probs = c(0.05, 0.95))

plot_counts_per_construct <- ggplot(counts_per_construct, aes(x = rank, y = reads_count)) +
  geom_point() +
  theme_classic() +
  scale_y_log10(breaks = c(quantiles[1], quantiles[2]), 
                     labels = c(quantiles[1], quantiles[2])) +
  geom_hline(yintercept = quantiles[1], linetype = "dashed") +
  geom_hline(yintercept = quantiles[2], linetype = "dashed") +  
  annotate("text", x = 25000, y = 5, label = paste0("FC = ", round(quantiles[[2]]/quantiles[[1]], 2)), size = 4) +
  annotate("text", x = 25000, y = 3, label = paste0("Mean = ", round(mean(counts_per_construct$reads_count), 2)), size = 4) +
  labs(title = "Reads per unique construct",
       x = "Uinque construct",
       y = "Read counts") +
  theme_minimal()

barcodes_per_construct <-filtered %>%
  group_by(construct) %>%
  summarise(barcodes_count = n_distinct(BARCODE)) %>%
  mutate(rank = rank(barcodes_count)) %>%
  arrange(rank)

# Calculate quantiles for barcodes
quantiles <- quantile(barcodes_per_construct$barcodes_count, probs = c(0.05, 0.95))

plot_barcodes_per_construct <- ggplot(barcodes_per_construct, aes(x = rank, y = barcodes_count)) +
  geom_point() +
  theme_classic() +
  scale_y_log10(breaks = c(quantiles[1], quantiles[2]), 
                     labels = c(quantiles[1], quantiles[2])) +
  geom_hline(yintercept = quantiles[1], linetype = "dashed") +
  geom_hline(yintercept = quantiles[2], linetype = "dashed") +  
  annotate("text", x = 25000, y = 3, label = paste0("FC = ", round(quantiles[[2]]/quantiles[[1]], 2)), size = 4) +
  labs(title = "Barcodes per unique construct",
       x = "Uinque construct",
       y = "Barcode counts") +
  theme_minimal()

plot_counts_per_construct + plot_barcodes_per_construct

```

### Barcode counts per cponstruct and read counts per construct are correlated

```{r another, include = TRUE}
counts_per_construct_not_ranked <-filtered %>%
  group_by(construct) %>%
  summarise(reads_count = sum(READS))

barcodes_per_construct_not_ranked <-filtered %>%
  group_by(construct) %>%
  summarise(barcodes_count = n_distinct(BARCODE))

data <- data.frame(
  Reads= counts_per_construct_not_ranked$reads_count,
  Barcodes = barcodes_per_construct_not_ranked$barcodes_count
)

# Plot
ggplot(data, aes(x = Reads, y = Barcodes)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Correlation between barcodes and reads counts")
```

## Constructs per Barcode?
We want to be able to identify which unique barcode corresponds to which construct, so non-unique barcodes will be eliminated. 

```{r barcodes duplicates, include = TRUE}
check <- filtered %>%
  group_by(BARCODE) %>%
  summarise(count = n_distinct(construct)) %>%
  arrange(desc(count))

print(head(check))
print(paste0(sum(check$count > 1), " = ", as.integer((sum(check$count > 1)/nrow(check))*100), "% out of ", nrow(check), " barcodes are combined with more than one construct"))
```


