---
title: "Barcode Association Report"
author: "Julia Rühle/Lera Dragan"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    code_folding: hide
    toc_depth: 4
    toc_float:
      collapsed: no
    highlight: pygments
  pdf_document:
    toc: yes
    toc_depth: '4'

  editor_options:
  chunk_output_type: console
# params:
#   data_file: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_knit$set(root.dir = "..")

require(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)

# Read the csv file and add 2 columns separating 2 gRNA from the GFP-barcode
# all <- read.csv(gzfile("C:/Users/Lera/Downloads/counts_matrix_20240905_134258.csv.gz"), sep = "\t")
all <- read.csv(gzfile(params$data_file), sep = "\t")
all$exp <- "gib2"

# Added a step to separate guide "barcode" from the GFP barcode, only works if the barcode line has | in the middle of it.
all <- all %>%
  separate(BARCODE, into = c("BARCODE", "GUIDE"), sep = "\\|")

# Filter out lines that were not mapped with bwa to GRE or gRNA
mapped <- all %>%
  filter(CRS != "None", GUIDE != "") %>%
  mutate(construct = paste0(CRS, GUIDE))
```

# Mapping Stats

```{r mapping stats}
tot_reads <- sum(all$READS)

perc_mapped <- sum(mapped$READS)/sum(all$READS)


mapping_stat <- data.frame(
  Library = "Gib2",
  "Total No of Reads" = tot_reads, 
  "Mapping Rate total" = perc_mapped
)

knitr::kable(mapping_stat, align = 'c')
```

# Barcode representation
Most importantly, we want to make sure that unique constructs are represented evenly in the library. We would expect equal number of reads/barcodes for each construct after sequencing. To look into this, we first compute the number of reads/barcodes per construct. Then we rank each construct, according to its number of reads (Rank 1: The construct with the lowest number of reads).

## How many READ COUNTS represent each construct? 

Each dot represents a unique construct ranked according to the number of reads by which it is represented. 

```{r number of reads representing constructs, include = TRUE}
counts_per_construct <-mapped %>%
  group_by(construct) %>%
  summarise(reads_count = sum(READS)) %>%
  mutate(rank = rank(reads_count)) %>%
  arrange(rank)

quantiles_reads <- quantile(counts_per_construct$reads_count, probs = c(0.05, 0.95))

ggplot(counts_per_construct, aes(x = rank, y = reads_count)) +
  geom_point() +
  theme_classic() +
  scale_y_log10(breaks = c(quantiles_reads[1], quantiles_reads[2]), 
                     labels = c(quantiles_reads[1], quantiles_reads[2])) +
  geom_hline(yintercept = quantiles_reads[1], linetype = "dashed") +
  geom_hline(yintercept = quantiles_reads[2], linetype = "dashed") + 
  labs(title = "Reads per unique construct",
       x = "Unique construct",
       y = "Read counts") +
  theme_minimal()

```

The mean number of reads per unique construct is **`r round(mean(counts_per_construct$reads_count), 2)`**.

Fold change between the 95% and 5% quantiles is **`r round(quantiles_reads[[2]]/quantiles_reads[[1]], 2)`**.

## How many BARCODES represent each construct? 

Each dot represents a unique construct ranked according to the number of barcodes by which it is represented. 

```{r number of barcodes representing constructs, include = TRUE}
barcodes_per_construct <-mapped %>%
  group_by(construct) %>%
  summarise(barcodes_count = n_distinct(BARCODE)) %>%
  mutate(rank = rank(barcodes_count)) %>%
  arrange(rank)

quantiles_bc <- quantile(barcodes_per_construct$barcodes_count, probs = c(0.05, 0.95))

ggplot(barcodes_per_construct, aes(x = rank, y = barcodes_count)) +
  geom_point() +
  theme_classic() +
  scale_y_log10(breaks = c(quantiles_bc[1], quantiles_bc[2]), 
                     labels = c(quantiles_bc[1], quantiles_bc[2])) +
  geom_hline(yintercept = quantiles_bc[1], linetype = "dashed") +
  geom_hline(yintercept = quantiles_bc[2], linetype = "dashed") +  
  labs(title = "Barcodes per unique construct",
       x = "Unique construct",
       y = "Barcode counts") +
  theme_minimal()

```
The mean number of barcodes per unique construct is **`r round(mean(counts_per_construct$reads_count), 2)`**.

Fold change between the 95% and 5% quantiles is **`r round(quantiles_bc[[2]]/quantiles_bc[[1]], 2)`**.

## Are barcode counts per construct and read counts per construct correlated?

```{r another, include = TRUE}
counts_per_construct_not_ranked <-mapped %>%
  group_by(construct) %>%
  summarise(reads_count = sum(READS))

barcodes_per_construct_not_ranked <-mapped %>%
  group_by(construct) %>%
  summarise(barcodes_count = n_distinct(BARCODE))

data <- data.frame(
  Reads= counts_per_construct_not_ranked$reads_count,
  Barcodes = barcodes_per_construct_not_ranked$barcodes_count
)

ggplot(data, aes(x = Reads, y = Barcodes)) +
  geom_point() +
  geom_smooth(formula = 'y ~ x', method = "lm", se = TRUE) +
  labs(title = "Correlation between barcodes and reads counts")
```

## Constructs per Barcode?

We want to be able to identify which unique barcode corresponds to which construct, so non-unique barcodes will be eliminated. 

```{r barcodes duplicates, include = TRUE}
check <- mapped %>%
  group_by(BARCODE) %>%
  summarise(count = n_distinct(construct)) %>%
  arrange(desc(count))

print(paste0(sum(check$count > 1), " = ", as.integer((sum(check$count > 1)/nrow(check))*100), "% out of ", nrow(check), " barcodes are combined with more than one construct"))
```
